\paperTitle

### 2014-01-31 doc/icml2014/directed.tex
Percy's changes

\section{Consistent Parameter Estimation in Directed Models}
\label{sec:directed}

The tensor factorization method attacks the heart of the non-convexity
  in latent variable models.
  Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  | h_j)$ for some $v, j$, we present a systematic approach to learn the
  conditional probability tables for every clique for a class of directed
  graphical models.
We begin by illustrating our approach with an example, followed by
  describing the algorithm in full generality.
For ease of exposition, we will describe our algorithm entirely in the
  context of directed graphical models, though it generalizes to factor
  graphs.
We also present the algorithm solely in terms of the method of moments; in
  \sectionref{piecewise}, we show how we can get a statistically more
  efficient estimator using convex optimization.

\subsection{Example: Directed grid model}

@@

Consider the directed grid model shown in \figureref{grid} which
  captures all the different dependency structures possible in
  a Bayesian network.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \cdots, h_4$.
The parameters of this model are the conditional probability tables
  $\pi \eqdef \Pr(h_1), T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1),
  V \eqdef \Pr(h_4 | h_2, h_3)$ and $O \eqdef \Pr(x^a_i | h_i)
  =  \Pr(x^b_i | h_i)$. 
Finally, we assume $O$ and $T$ have full column rank.

\paragraph{Estimating $O$}
Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, allowing us to use
  \TensorFactorize, as reviewed in \sectionref{setup}, to recover
  the $O$.

\paragraph{Estimating $\pi$}
The moments of $x^a_1$,$\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$; $\mO_1 = O \pi$. 
By \assumptionref{full-rank}, $O$ has full column rank and thus can be
  inverted to recover $\pi$; $\pi = \pinv O \mO_1$.

\paragraph{Estimating $T$}
Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
  \eqdef \Pr(x^a_1, x^a_2)$ in terms of the pairwise marginals $\mH_{12}
  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$;
\begin{align*}
  \mO_{12} &= O \mH_{12} O^T \\
  \mH_{12} &= \pinv O \mO_{12} \pinvt O.
\end{align*}
$T$ can be recovered from the $\mH_{12}$ by suitable normalization.

\paragraph{Estimating $V$}
Finally, we can estimate $V$ from the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$ as
follows,
\begin{align*}
  \mO_{234} &= \mH_{234}(O, O, O) \\
  \mH_{234} &= \mO_{234}(\pinv O, \pinv O, \pinv O).
\end{align*}

\subsection{General algorithm}

Fundamentally, our algorithm takes advantage of having a view for every
  hidden variable in a clique to help identify its hidden state. We define this condition as follows,
\begin{property}(Exclusive views)
  \label{prop:exclusive-views}
A clique $\sC$ of $\sG$ is said to posses exclusive views if for every hidden variable $h_i \in \sC$,
  there is some observed variable, $x_{v}$, which is conditionally
  independent of the rest of the clique given $h_i$ and whose conditional
  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} | h_i)$ is known. 
We call $x_v$ the exclusive view of $h_i$ in $\sC$.
\end{property}

@@

We now show this condition is sufficient to learn the marginal
  distribution of the clique.
Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
  $x_{v_j}$ be the exclusive view for $h_{i_j}$ in $\sC$ and $\sV
  = \{x_{v_1}, \cdots, x_{v_m}\}$ be the set of views for the clique
  $\sC$.
We can write down the moments $\mO_\sV$,
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh \in \sH}
      \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{\vh \in sH} \Pr(\Sh{\sC}) 
          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
hidden marginals $Z_\sC$ by inverting the conditional moments,
\begin{align*}
  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}).
\end{align*}
Finally, the conditional probability tables for $\sC$ can easily be gotten from
  $Z_\sC$ via conditioning.
\algorithmref{learnclique} summarizes the procedure, \LearnClique.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnClique}
  \label{algo:learnclique}
  \begin{algorithmic}
    \REQUIRE A $\sC$ with \propertyref{exclusive-views}.
    \ENSURE The marginal distribution of the clique, $Z_\sC$.
      \STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
      \STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
  \end{algorithmic}
\end{algorithm}

@@

If we can guarantee that every clique in a graph possesses exclusive
  views, then we can apply \LearnClique iteratively to learn the entire
  model.
In the example, we were able to guarantee \propertyref{exclusive-views}
  by identifying a hidden variable, $h^1$, with three conditionally
  independent observed variables, $x^a_1, x^b_1, x^a_2$, followed by
  using \TensorFactorize to learn the conditional moments.
We define such a hidden variable to be a bottleneck;
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
    least three conditionally independent observed variables, or views,
    $x_1, x_2, x_3$. 
  Let $\sV_{h}$ be the set of views for $h$.
\end{definition}
Finally, the claim we make is that we can recover parameters for any
  graphical model where every hidden variable is a bottleneck;
\begin{property}
  \label{prop:bottleneck}
  Every hidden variable is a bottleneck.
\end{property}
We note that \propertyref{bottleneck} can be relaxed if some cliques
  share parameters;
For example, because we can identify the observation potential from just
  $x^a_1, x^b_1$ and $x^a_2$, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can correspondingly omit the observations $x^b_2,
  x^b_3$ and $x^b_4$.
In this case, we only require that the hidden variables in distinct
  cliques share parameters.

Applying \TensorFactorize to every bottleneck gives us candidate views
  for every hidden variable, provided \assumptionref{full-rank} holds for each bottleneck;
\begin{assumption} 
  \label{asm:full-rank-plus}
  For every hidden variable $h_i$, $\pi\oft{i} \succ 0$ and
    $\mOpp{v}{i}$ has full column rank.
\end{assumption}

Next, we must show that these candidate views actually imply that
  \propertyref{exclusive-views} holds for every clique in the graph.
Before we establish this condition, we describe a reduction to
  a canonical form that will allow us to handle the case where an observed
  variable has multiple parents.

\begin{lemma}(Canonical form) 
  \label{lem:reduction}
Every directed graphical model can be transformed into one in which
  the observed variables are leaves with exactly one parent. 
There is a one-to-one correspondence between the parameters of this
  transformed model and the original one.
\end{lemma}
\begin{proof}
  \begin{figure}
    \centering
    \subimport{figures/}{reduction.tikz}
    \caption{Reduction to canonical form.}
    \label{fig:reduction}
  \end{figure}

  %\providecommand{\hp}{\ensuremath{h_{\textrm{phantom}}}}
  \providecommand{\hp}{\ensuremath{h_p}}

  Let $x_v$ be an observed variable with parents $P$ and children $C$.
  Consider the following transformation.
  Replace $x_v$ with a phantom hidden variable \hp with the same 
  domain\footnote{Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      arbitrary sizes, provided \assumptionref{full-rank-plus} holds.},
  parents $P$ and children $C \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). Also, set $\mOpp{v}{p} = I$.

@@

\begin{lemma}(Bottlenecks guarantee exclusive views)
  Provided \propertyref{bottleneck} holds, every hidden variable
    $h_{i_1}, \cdots, h_{i_m}$ in a clique $\sC$ has a exclusive view.
\end{lemma}
\begin{proof}
  By \lemmaref{reduction}, we can assume w.l.o.g. that every observed
  variable is a leaf in the graph, causing any v-structures in the graph
  to be active. This allows us to reason about independence simply by
  checking whether or not two nodes are connected or not; note that this
  proof applies for undirected graphs as well.

  Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
    views for it and let $\sC^- \eqdef \sC \setminus \{h\}$.
  By definition, the views of $h$ are conditionally independent, viz.
    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
    $h$.
  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
  \in \sC^-$

@@

@@

      \FOR{$h_i \in H$} 
        \STATE Apply \TensorFactorize to learn conditional moments
        $\mOpp{v}{i}$ for every $\sV_h$.

@@

@@

\LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
  \LearnClique, and is thus consistent itself. 
We next describe the asymptotic variance for \LearnClique when using the
  method of moments.
\begin{lemma}(Asymptotic variance of $Z_\sC$)
  \label{lem:mom-variance}  
  The asymptotic variance of $Z_\sC$, $\Sigma^\mom_\sC$ is,
  \begin{align*}
    \Sigma^{(\mom)} &= \mOppit{\sV}{\sC} \Sigma_\sV \mOppi{\sV}{\sC},
  \end{align*}
  where $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
    \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $k^m \times d^m$ matrix, and 
\begin{align*}
  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
\end{align*}
  is the variance of the observations $M_\sV$.
\end{lemma}
\begin{proof}
We have for each clique,
\begin{align*}
  Z_{\sC} &= \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m}) \\
\end{align*}

Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
matrix, we can rewrite this representation as,
\begin{align*}
  Z_{\sC} &= \pinv{\mOpp{\sV}{\sC}} \mO_\sV.
\end{align*}

We know the asymptotic distribution of $M_\sV$ goes as,
\begin{align*}
  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
\end{align*}
where $\Sigma_\sV$ is the variance of the observations. 
Finally, we have that $Z_\sC$ goes as
\begin{align*}
  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
\end{align*}
\end{proof}

Note that extending these results to finite sample bounds can be done
  via a straightforward application of perturbation bounds.

>>>>

% DONE: too wordy
%\section{Consistent Parameter Estimation in Directed Models}
\section{Directed models}
\label{sec:directed}

[TODO: need to say we're going to do this for the infinite data case,
and then study bounds]

[TODO: try to be consistent: either use latent or hidden]

% DONE: need more setup of what the goal is
%The tensor factorization method attacks the heart of the non-convexity
  %in latent variable models.
  In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
  In directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
  are only a subset of the parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
  Let us see how we can estimate the remaining parameters with an example (Section~\ref{sec:directedExample}).
  Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
For ease of exposition, we will describe our algorithm entirely in the
  context of directed graphical models, though it generalizes to factor
  graphs.
We also present the algorithm solely in terms of the method of moments; in
  \sectionref{piecewise}, we show how we can get a statistically more
  efficient estimator using convex optimization.

\subsection{Example: Directed grid model}
\label{sec:directedExample}

@@

Consider the directed grid model shown in \figureref{grid} which
  % DONE: all the different dependency => too strong
  captures the local dependency structures possible in
  a Bayesian network.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \cdots, h_4$.
The parameters of this model are the conditional probability tables
$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$. 
Finally, we assume $O$ and $T$ have full column rank.

\paragraph{Estimating $O$}
Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, allowing us to use
  \TensorFactorize, as reviewed in \sectionref{setup}, to recover
  $O$.

\paragraph{Estimating $\pi$}
The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$ by $\mO_1 = O \pi$. 
By \assumptionref{full-rank}, $O$ has full column rank and thus can be
  inverted to recover $\pi$: $\pi = \pinv O \mO_1$.

\paragraph{Estimating $T$}
Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
  \eqdef \Pr(x^a_1, x^a_2)$, in terms of the latent marginals $\mH_{12}
  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$:
\begin{align*}
\mO_{12} = O \mH_{12} O^\top \quad\Rightarrow\quad
  \mH_{12} = \pinv O \mO_{12} \pinvt O.
\end{align*}
$T$ can be recovered from the $\mH_{12}$ by suitable renormalization.

\paragraph{Estimating $V$}
Finally, we can estimate $V$ by renormalizing the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$,
which can be obtained analogously:
\begin{align*}
  \mO_{234} = \mH_{234}(O, O, O) \quad\Rightarrow\quad
  \mH_{234} = \mO_{234}(\pinv O, \pinv O, \pinv O).
\end{align*}

\subsection{General algorithm}
\label{sec:directedGeneral}

Note that in each of the above cases,
each hidden variable $h_i$ of a clique has its own view which is conditionally independent
given the other variables in the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
We capture this general property as follows:
\begin{property}(Exclusive views)
  \label{prop:exclusive-views}
  A clique $\sC$ of $\sG$ is said to possess \textbf{exclusive views property} if for every hidden variable $h_i \in \sC$,
  there is some observed variable, $x_{v}$, which is conditionally
  independent of the rest of the clique given $h_i$.
  and whose conditional
  moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ the exclusive view of $h_i$ in $\sC$.
\end{property}

@@

We now show this condition is sufficient to learn the marginal
  distribution of the clique.
Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $\sC$ (not necessarily unique) and $\sV
  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
We can write down the moments $\mO_\sV$ as follows:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh \in \sH}
      \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
hidden marginals $Z_\sC$ by inverting:
\begin{align*}
  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}).
\end{align*}
Finally, the conditional probability tables for $\sC$ can easily be gotten from
  $Z_\sC$ via renormalization.
\algorithmref{learnclique} summarizes the procedure, \LearnClique.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnClique}
  \label{algo:learnclique}
  \begin{algorithmic}
    \REQUIRE Clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
    \ENSURE Marginal distribution of the clique $Z_\sC$.
      \STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
      \STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
  \end{algorithmic}
\end{algorithm}

@@

% DONE: transition the content more gradually (checkpoint: what have we done so far, what next)
% DONE: refer to properties by name in addition to by number
We have shown how to estimate the parameters of any clique that possesses the exclusive
  views property (\propertyref{exclusive-views}).  But which cliques have this property?
In the example above, we were able to guarantee \propertyref{exclusive-views}
  by identifying a hidden variable, $h^1$ along with three conditionally
  independent observed variables, $x^a_1, x^b_1, x^a_2$.
  %followed by using \TensorFactorize to learn the conditional moments.

We define such a hidden variable to be a bottleneck:
\begin{definition}[Bottleneck]
  A hidden variable $h$ is said to be a \textbf{bottleneck} if there exists at
    least three conditionally independent observed variables, or views,
    $x_1, x_2, x_3$. 
  Let $\sV_{h}$ denote a set of views for $h$ (not necessarily unique).
\end{definition}

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
We will prove that the following simple property guarantees that all cliques
have the exclusive views property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  Every hidden variable is a bottleneck.
\end{property}

%Applying \TensorFactorize to every bottleneck gives us a set of views
%  for every hidden variable. 
%To proceed, we need to check that \
%%(i) understand the assumptions under which
%  %\assumptionref{full-rank} holds for each bottleneck, and (ii)
%  that the candidate views actually imply that \propertyref{exclusive-views}
%  holds for every clique in the graph.

% PL: this is not obviously correct
%To address the first, we will need that the following assumption on the
%  parameters of the model:
%\begin{assumption} 
%  \label{asm:full-rank-plus}
%  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, let
%    $Z_\sC$ be the marginal distribution for that clique. 
%  Then every mode-$i$ unfolding of $Z_\sC$ has full column rank.\verify
%\end{assumption}
%Intuitively, the assumption guarantees that the marginal probability of
%  any hidden state is non-zero.

Before we establish this relationship,
we describe a reduction of any graphical model to
  a canonical form to make our arguments simpler.
  % DONE: don't say this yet since reader can't appreciate it.
  %that will allow us to handle the case where an observed
  %variable has multiple parents.

\begin{lemma}[Canonical form]
  \label{lem:reduction}
Every directed graphical model can be transformed into one in which
  the observed variables are leaves with exactly one parent. 
There is a one-to-one correspondence between the parameters of this
  transformed model and the original one.
\end{lemma}
\begin{proof}
  \begin{figure}
    \centering
    \subimport{figures/}{reduction.tikz}
    \caption{Reduction to canonical form.}
    \label{fig:reduction}
  \end{figure}

  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}

  Let $x_v$ be an observed variable with parents $P$ and children $C$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). 
  The domain of $\hp$ is the same as $x$,
    and $\mOpp{v}{\textrm{new}} = I$.\footnote{Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      arbitrary sizes.
      [TODO: add qualifier]
      %provided \assumptionref{full-rank-plus} holds.
      }

@@

\begin{lemma}[Bottlenecks guarantee exclusive views]
  Provided \propertyref{bottleneck} holds, every hidden variable
    $h_{i_1}, \cdots, h_{i_m}$ in a clique $\sC$ has a exclusive view.
\end{lemma}
\begin{proof}
  By \lemmaref{reduction}, we can assume that every observed
  variable is a leaf in the graph, causing any v-structures in the graph
  to be active. This allows us to reason about independence simply by
  checking whether or not two nodes are connected or not. Note that this
  proof applies for undirected graphs as well.

  Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
    views for it and let $\sC^- \eqdef \sC \setminus \{h\}$.
  By definition, the views of $h$ are conditionally independent, viz.
    every connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
    $h$.
  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
  \in \sC^-$.

@@

[TODO: somewhere give examples that are not in our class, preferably the simplest one, cite Sontag]

@@

      \FOR{each hidden variable $h_i \in H$} 
        \STATE Apply \TensorFactorize to learn conditional moments
        $\mOpp{v}{i}$ for every $\sV_{h_i}$.

@@

% DONE: move here because not central to story
\paragraph{Remark.} We note that \propertyref{bottleneck} can be relaxed if some cliques
  share parameters.
In the directed grid model, we can identify the conditional moments $O$ from just
  $x^a_1, x^b_1$ and $x^a_2$.
  Therefore, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.
  [TODO: make this a footnote if running out of space]
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.

@@

\LearnMarginals~ combines two consistent algorithms, \TensorFactorize and
\LearnClique, [TODO: too much space after LearnClique] and is thus consistent itself.
[TODO: did they prove this; if so just say "they show", not "we can show"]
Using results from
  \citet{anandkumar12moments,anandkumar13tensor} we can show that
  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
  x_{v_2}, x_{v_3}$ has the following sample complexity,
  [TODO: define hat O]
  [TODO: max subscript is misaligned]
\begin{align*}
  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
\end{align*}

[TODO: call MOM squared]
We next describe the asymptotic variance for \LearnClique when using the
  method of moments.
\begin{lemma}(Asymptotic variance of $Z_\sC$)
  \label{lem:mom-variance}  
  The asymptotic variance of $Z_\sC$ is:
  \begin{align*}
    \Sigma^{\mom} &= \mOppAlli \Sigma_\sV \mOppAllit,
  \end{align*}
  where $\pinv{\mOppAll} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes
  \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$, a $d^m \times k^m$ matrix, and 
\begin{align*}
  \Sigma_\sV &= \text{\rm diag}(M_\sV) - M_\sV M_\sV^\top,
\end{align*}
  is the asymptotic variance of the observations $M_\sV$.
\end{lemma}
\begin{proof}
We have for each clique,
\begin{align*}
  Z_{\sC} &= \mO_\sV(\mOppi{v_1}{i_1},\cdots,\mOppi{v_m}{i_m}) \\
\end{align*}

%Choosing to represent $Z_\sC$ and $M_\sV$ as vectors and
%$\mOppi{v_1}{i_1} \otimes \cdots \otimes \mOppi{v_m}{i_m}$ as the
%matrix,
We can rewrite this representation as:
\begin{align*}
  Z_{\sC} &= \pinv{\mOppAll} \mO_\sV.
\end{align*}

From classical asymptotics \cite{vaart98asymptotic},
we have that the asymptotic variance of $M_\sV$ is:
\begin{align*}
  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
\end{align*}
where $\Sigma_\sV$ is the variance of the observations. 
Finally, we have that $Z_\sC$ goes as
\begin{align*}
  \sqrt{n}(\hat Z_\sC - Z_\sC) \convind \sN(0, \pinvt{\mOppAll} \Sigma_\sV \pinv{\mOppAll}).
\end{align*}
\end{proof}

% DONE: too brazen to say
%Note that extending these results to finite sample bounds can be done
  %via a straightforward application of perturbation bounds.

### 2014-01-31 doc/icml2014/discussion.tex
Percy's changes

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar12linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
Thus, while \propertyref{bottleneck} is much stronger than
  \propertyref{exclusive-views}, we expect that without any knowledge of
  the correlation between variables, \propertyref{bottleneck} is
  necessary for identifiability.

>>>>

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
Thus, while \propertyref{bottleneck} is much stronger than
  \propertyref{exclusive-views}, we expect that without any knowledge of
  the correlation between variables, \propertyref{bottleneck} is
  necessary for identifiability.
